---
title: "Week 1: Intro to dplyr"
author: "Hazel KAVILI"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Install Libraries

```{r, warning=FALSE, message=FALSE, echo=TRUE}
#find mine's blog post where explain why directly load tidyverse instead of dplyr
library(tidyverse)
```

## Load Data Into R

```{r, warning=FALSE, message=FALSE, echo=TRUE}
##don't forget to add the link data to read from csv and mention that you can read it from web or from your pc
nordic_data <- read_csv("nordic_data.csv")
```

## Summary of data
There are some functions to check the summary of your data: `str`, `summary`, `glimpse` 

```{r, warning=FALSE, message=FALSE}
glimpse(nordic_data)
```

## Comparison with base R and Tidyverse
```{r, warning=FALSE, message=FALSE}

```

## Did you see *Pipes*? *( %>% )*
```{r, eval = FALSE, warning=FALSE, message=FALSE, echo = TRUE }
nordic_data %>% 
  select(Country_Name, Fertility_rate_total_births_per_woman_, year) %>% 
  filter( year == 1960) 
```

## *Select:* Choosing is not losing!
```{r, warning=FALSE, message=FALSE}

```

## *Filter:* Do we want everything?
```{r, warning=FALSE, message=FALSE}

```

## How should we fill the blank areas?
```{r, eval=FALSE}

```

## Here is the solution:

```{r}
```


## *Mutate:* I need a new column!
```{r, warning=FALSE, message=FALSE}

```

## *Arrange:*  Let's order the rows
```{r}

```

## Let's put it all together!
```{r}

```

## Here is the solution
```{r}
```




## Good job!

Thank you for joining the training! 


