---
title: "Week 2: Intro to Data Wrangling with dplyr"
author: "Hazel KAVILI"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: flatly
---

## Install Libraries
We'll load the packages with `library(tidyverse)` instead of loading the core packages one by one. 
Today we'll focus on `dplyr`, a package that transforms data. dplyr implements a grammar for transforming tabular data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

## Load Data Into R
Next thing we'll do is load the data into our R environment. We'll use a data set about coffee ratings from weekly [Tidy Tuesday challenge ](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md). 

We'll use `read_csv` function from `readr` package to get the data, and assign it to `coffee_ratings`. In the rest of the document, we'll use that name to call our data frame.

```{r, warning=FALSE, message=FALSE, echo=TRUE}
coffee_ratings <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
```

## Summary of data
Let's look at our data frame to get to know it.
There are some functions to check the structure and summary of the data: `str`, `summary`, `glimpse` are the most common ones.

```{r, warning=FALSE, message=FALSE, eval=FALSE, echo=TRUE}
glimpse(coffee_ratings)
```

```{r, warning=FALSE, message=FALSE}
head(coffee_ratings, n = 3)
```

There are 43 variables (columns) and 1339 rows in our data frame. There is data for both Arabica and Robusta beans, across many countries and professionally rated on a 0-100 scale. All sorts of scoring/ratings for things like acidity, sweetness, fragrance, balance, etc - may be useful for either separating into visualizations/categories or for modeling/recommenders.

## Questions arise
- Which coffee species has the highest total_cup_points? 
- What is the average total_cup_points for each species?
- How many species are there for each country_of_origin?

## Comparison with base R and Tidyverse
Before moving into dplyr verbs, I'd like to show you a comparison of base R and tidyverse syntax. 
The two code chunks are doing the same thing: selecting two columns from our dataframe. 

```{r, warning=FALSE, message=FALSE, eval=FALSE, echo=TRUE}
coffee_ratings[ , c('species', 'country_of_origin')]
```

```{r, warning=FALSE, message=FALSE, eval=FALSE, echo=TRUE}
  select(coffee_ratings, species, country_of_origin)
```


## *Select:* Choosing is not losing!
We are moving to our first dplyr verb. `select` helps us select the variables we want to use or we need for our analysis. 

```{r, warning=FALSE, message=FALSE, eval=TRUE, echo=TRUE}
 select(coffee_ratings, species, country_of_origin, color, certification_body) 
```

`select` has helpers like the one on the below. You can select range of columns by using `:`

```{r, warning=FALSE, message=FALSE}
  select(coffee_ratings, 1:4, 21:34) 
```


## *Filter:* Do we want everything?
`filter` helps us to extract rows that meet logical criteria. 
Let's say we want to look at our coffee_ratings dataset only for Ethiopia originated coffees. Then what we need to do is `filter` our dataset by using necessary column and value:

```{r, warning=FALSE, message=FALSE, eval=TRUE, echo=TRUE}
filter(coffee_ratings, country_of_origin == "Ethiopia")
```

```{r, warning=FALSE, message=FALSE, eval=TRUE, echo=TRUE}
filter(coffee_ratings, number_of_bags >= 60)
```

## *Arrange:*  Let's bring some order to data!
`arrange` order rows from smallest to largest values. 
Let's say we want to look at our data frame arranged by `total_cup_points`. In default, your dataset will be arranged bu ascending order.

```{r, eval=TRUE, echo=TRUE}
  arrange(coffee_ratings, total_cup_points)
```

If we want to order it by descending order of total_cup_points, we can use this syntax.
```{r, eval=TRUE, echo=TRUE}
  arrange(coffee_ratings, desc(total_cup_points))
```

## Steps

```{r, eval=FALSE}
coffee_ratings_ethiopia <- select(coffee_ratings, total_cup_points, species, country_of_origin)
coffee_ratings_ethiopia <- filter(coffee_ratings_ethiopia, country_of_origin == "Ethiopia")
coffee_ratings_ethiopia <- arrange(coffee_ratings_ethiopia, total_cup_points)
coffee_ratings_ethiopia
```

##  *Pipes* *( %>% )*
Pipe helps us to connect our dplyr verbs. 
Passes results on left into first argument of function on right. 

Let's try to do the same thing with pipes and read the code below: 
I have the dataframe called coffee_ratings, I call it. 
I select 3 variables from that data frame and filter it by country_of_origin variable equals to Ethiopia. Then I arrange it by total_cup_points variable in ascending order.


```{r, eval = FALSE, warning=FALSE, message=FALSE, echo = TRUE}
coffee_ratings_ethiopia <- 
  coffee_ratings %>%
  select(species, certification_body, country_of_origin) %>% 
  filter(country_of_origin == "Ethiopia") %>% 
  arrange(total_cup_points)
```

What we see in here is that everything is chained. I can read that code from left to right like a sentence.

## Structure of dplyr functions
Most of the dplyr verbs have that structure:
```{r, eval=FALSE, echo=TRUE}
1- dplyr_verb(data_frame, colum_name, some_other_arguments)

2- data_frame %>% 
    dplyr_verb(column_name, some_other_arguments) %>% 
    dplyr_verb(column_name, some_other_arguments)
```


## Your Turn: How should we fill the blank areas?
Now it's time for a pop-quiz! 
We need to:
- Select **total_cup_points, species, owner, countyr_of_origin and number_of_bags** 
- Then Filter **country of origin for only Brazil** and **total cup points is higher than 70**
- Then Arrange it by number_of_bags in descending order.

```{r, eval=FALSE, echo = TRUE}
coffee_ratings data %>% 
  ____(_:_, number_of_bags) %>% 
  ____(____ == "Brazil") %>%
  ____(total_cup_points > ___) %>% 
  ____(desc(number_of_bags))
```

## Here is the solution:
```{r, eval=TRUE, echo = TRUE}
coffee_ratings %>% 
  select(1:4, number_of_bags) %>% 
  filter(country_of_origin == "Brazil") %>%
  filter(total_cup_points > 70) %>% 
  arrange(desc(number_of_bags))
```


## *Group_by & Summarise:* Let's create associations among groups!

We often want to look at our data by some group (if we have) for example country, year, age, gender, education, etc. to spot some differences (if there is).
`group_by` gives the data frame a grouping using one or more columns, which modifies the subsequent call to `summarise`. 

We can use functions like `mean`, `median`, `sum`, `sd`, `max`, `min` to find out some group statistics inside of `summarise` function.

```{r, eval=TRUE, echo = TRUE, warning=FALSE, message=FALSE}
coffee_ratings %>%
  group_by(species) %>%
  summarise(avg_rating = mean(total_cup_points),
            proportion = n()/nrow(coffee_ratings))
```


```{r, eval=TRUE, echo = TRUE, warning=FALSE, message=FALSE}
coffee_ratings %>%
  group_by(country_of_origin) %>%
  summarise(avg_rating = mean(total_cup_points),
            median_rating = median(total_cup_points)
            )
```

## *Mutate:* We need a new column!

`mutate` lets us add or overwrite columns by computing a new value for them.
For example, in the code below, we want a column name `is_processing_method_full` which is filled by checking if `processing_method` have some value or not. If it is not NA, we'll write 1, otherwise 0. 

We didn't have `is_processing_method_full` before but we create it by using `processing_method` variable.

```{r, eval=TRUE, echo = TRUE}
coffee_ratings %>% 
  select(species, country_of_origin, processing_method) %>% 
  mutate(is_processing_method_full = ifelse(!is.na(processing_method), 1, 0)) 
    
```


## Your Turn: Let's put it all together!
We want to look at the mean of aroma, flavor and aftertast variables by countries.
Can we find the true order of that code piece?

```{r, eval = FALSE, echo = TRUE}
1- filter(species == "Arabica")  
2- filter(!is.na(country_of_origin)) 
3- select(total_cup_points, species, country_of_origin, processing_method, aroma, flavor, aftertaste)
4- coffee_ratings
5- arrange(mean_aroma) 
6- summarise(mean_aroma = mean(aroma),
            mean_flavor = mean(flavor),
            mean_aftertaste = mean(aftertaste))
7- group_by(country_of_origin)

```


## Here is the solution:
```{r, eval = TRUE, echo = TRUE, warning=FALSE, message=FALSE}
coffee_ratings %>% 
  select(total_cup_points, species, country_of_origin, processing_method, aroma, flavor, aftertaste) %>%
  filter(species == "Arabica") %>% 
  filter(!is.na(country_of_origin)) %>% 
  group_by(country_of_origin) %>% 
  summarise(mean_aroma = mean(aroma),
            mean_flavor = mean(flavor),
            mean_aftertaste = mean(aftertaste)) %>%
  arrange(mean_aroma) 
```


## Good job!
Thank you for joining the training! 

You can find the all lesson materials my [Github page.](https://github.com/UniversalTourist/tidyverse_trainer_exam_materials)

Any questions? I'd be happy to answer! 




